---
title: "Diverse Cohorts RNA-Seq QC (MSSM)"
author: "Jaclyn Beck"
format:
  html:
    df-print: kable
    toc: true
    toc-depth: 2
editor: visual
code-fold: true
---

# Setup

<details>

```{r}
#| cache: false
#| label: read-chunk
#| include: false

knitr::read_chunk("qc_functions.R")

```

```{r}
#| label: include-libraries
#| output: false
```

```{r}
#| label: setup
#| output: false

dataset <- "MSSM"

# Setting to TRUE will upload the filtered counts file (post-QC) to the harmonization
# project on Synapse.
upload_to_synapse <- FALSE
```

## Load files

Metadata (individual, biospecimen, assay files)

```{r}
#| label: download-metadata
#| output: false
```

Counts file

```{r}
#| label: download-counts
#| output: false
```

FastQC and MultiQC stats

```{r}
#| label: download-qc-stats
```

</details>

# Validation

## FastQC

Examine Phred score and average base content at each position of a read.

Samples fail QC if they have a Phred score \< `{r} configs$thresholds$phred`.

Samples get a QC "warning" if their base content is above `Q3 + 1.5 * IQR` for any base or if their Phred score is below `Q1 - 1.5 * IQR`.

Samples with more than 2 warnings across all QC metrics fail QC.

```{r}
#| label: validate-fastqc
```

**FastQC failures / warnings:** `{r} n_fqc_warn_fail` sample(s)

```{r}
#| label: print-fastqc-results
```

## MultiQC

Examine percentage of reads mapped and percentage of reads duplicated.

Samples fail QC if they have \< `{r} thresholds$reads_mapped`% reads mapped or \> `{r} thresholds$reads_duplicated`% duplicated reads.

Samples get a QC warning if they have less than `Q1 - 1.5 * IQR` mapped reads or more than `Q3 + 3 * IQR` duplicated reads.

```{r}
#| label: validate-multiqc
```

**MultiQC failures / warnings:** `{r} n_mqc_warn_fail` sample(s)

```{r}
#| label: print-multiqc-results
```

## Inferred sex

Examine average expression of key Y-chromosome genes to infer the sex of each sample. Genes used are **RPS4Y1, EIF1AY, DDX3Y,** and **KDM5D,** which were determined to be robust markers for sex in [Staedtler, et al (2013)](https://www.ncbi.nlm.nih.gov/pubmed/23829492).

Samples whose inferred sex do not match their reported sex are marked as QC failures.

```{r}
#| label: validate-sex
```

**Reported sex vs inferred sex mismatches:** `{r} sum(!metadata$sex_valid)` sample(s)

```{r}
#| label: print-sex-mismatches
```

## PCA outliers

Samples were split by tissue and sequencing batch, PCA was run on each group, and outliers were defined as samples falling outside 4 standard deviations of the data along PC1 and PC2.

```{r}
#| label: validate-pca
```

**PCA outliers:** `{r} sum(!metadata$pca_valid)` sample(s)

```{r}
#| label: print-pca-outliers
```

## DV200

Samples were marked as QC failures if:

-   Both DV200 and RIN are `NA`

-   DV200 \< `{r} thresholds$DV200`

    -   or RIN \< `{r} thresholds$RIN`, if DV200 is `NA` for that sample

```{r}
#| label: validate-dv200
#| warning: false
```

**Samples with low DV200:** `{r} sum(!metadata$DV200_valid)` sample(s)

```{r}
#| label: print-dv200-results
```

# Save samples that passed QC

```{r}
#| label: save-samples
```

**`{r} ncol(counts)`** of **`{r} orig_size`** samples passed QC. Results by tissue:

```{r}
#| label: print-final-qc-results
```

#### Samples that will be removed

```{r}
#| label: print-qc-failures
```
